A finite-state morphological transducer for Kyrgyz

### 2011-10-24 dev/paper/abstract.tex
shortened more tables

In Kyrgyz, there are two basic archiphonemes: a low vowel, \{A\}, and a high vowel, \{I\}.  The high vowel takes the backness and rounding of the preceding vowel, resulting in the values shown in table~\ref{vowelharmonyI}.

\begin{table}[htbp]
	\centering
	\caption{vowel harmony for archiphoneme \{I\}}\label{vowelharmonyI}
	\subtable{
		\begin{tabular}{cc}
			\toprule
			after & result \\
			\midrule
			и & и \\
			ү & ү \\
			е & и \\
			ө & ү \\
			\bottomrule
		\end{tabular}
	}
	\hspace{2em}
	\subtable{
		\begin{tabular}{cc}
			\toprule
			after & result \\
			\midrule
			ы & ы \\
			у & у \\
			а & ы \\
			о & у \\
			\bottomrule
		\end{tabular}
	}
\end{table}

The back vowel \{A\} also takes the backness and rounding of the preceding vowel, with the exception of when it occurs after /у/---n this case, it has an unrounded variant, as depicted in table~\ref{vowelharmonyA}.

\begin{table}[htbp]
	\centering
	\caption{vowel harmony for archiphoneme \{A\}}\label{vowelharmonyA}
	\subtable{
		\begin{tabular}{ll}
			\toprule
			after & result \\
			\midrule
			и & е \\
			ү & ө \\
			е & е \\
			ө & ө \\
			\bottomrule
		\end{tabular}
	}
	%\qquad
	\hspace{2em}
	\subtable{
		\begin{tabular}{ll}
			\toprule
			after & result \\
			\midrule
			ы & а \\
			у & а \\
			а & а \\
			о & о \\
			\bottomrule
		\end{tabular}
	}
\end{table}

There were other vowel archiphonemes that had to be implemented in this project.  For example, \{U\} occurs in the past habitual suffix -/чU/ and the general gerund/infinitive -/Uː/, but nowhere else in the language.  Also, \{E\} was created for use in the habitual/future suffix; it has surface forms identical to those of \{A\}, except after vowels where it surfaces as [й].  Despite the fact that \{E\} is very similar to \{A\}, a separate but very similar twolc rule had to be created.  Repeated content could have been reduced by using cascading rules instead of two-level rules, but this would have caused other complications, such as finding the correct rule ordering.  An alternative possibility would’ve been to have a single intermediate level so that a rule like ``\{E\}→\{A\} after vowels'' with a default surface form of [й] could've been implemented.

>>>>

In Kyrgyz, there are generally two basic archiphonemes: a low vowel, \{A\}, and a high vowel, \{I\}.  The high vowel takes the backness and rounding of the preceding vowel, resulting in one of four surface vowels.  The back vowel \{A\} also takes the backness and rounding of the preceding vowel, with the exception of when it occurs after /у/---in this case, it has an unrounded variant, as depicted in table.
